The Galactic Midplane is Not a Plane: Implications for {\em Gaia}

### 2019-04-10 paper/ms.tex
Add keywords
Fix line indentation throughout

\shortauthors{Beane et al.}
\shorttitle{The dynamical solar neighborhood}

@@

\keywords{actions}

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a revolution in the
era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
major merger
\citep{2018ApJ...860L..11K,2018Natur.563...85H,2018arXiv180704290L,2019MNRAS.482.3426M},
a phase-space ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly indicating either local
substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L} or bar
buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation by a dark matter
subtructure in the tidal stream GD1 \citep{2018ApJ...863L..20P, 2018arXiv181103631B}.
These discoveries all indicate that the Milky Way's stellar distribution, which demonstrably departs significantly from axisymmetry, is undergoing phase mixing and dynamical interactions across a range of spatial and temporal scales.

@@

%In addition, the necessity of 6D phase space measurements of a star limits the Galactic volume observable in action space.

%With these limitations in mind, it is therefore natural to consider the reason to transform to action space.

@@

In Section~\ref{sec:discussion} we discuss the implications of midplane variations, and the resulting systematic uncertainty in the vertical action, for action-space analyses beyond the dynamical solar neighborhood. As an illustration, in Section \ref{sssec:neighborhood} we use these estimates to consider the feasibility of one science goal that relies on the invariance of actions --- the reconstruction of open clusters across large distances using ``dynamical tagging;'' that is, the association of stars to a birth cluster based on their similar actions. In light of our findings, we considered the orbits of the
nine open clusters within $250\,\pc$ as reported by the {\em Gaia}
collaboration \citep{2018AA...616A..10G}. In Table~\ref{tab:real_clusters}, we
report the distance to each cluster, the three actions, and the maximum
vertical extent of their orbits ($z_{\text{max}}$). 
%The actions are computed using the technique described in Section~\ref{ssec:action_comp}. 
 In Section~\ref{sssec:reconstruction} we estimate and discuss the size of the dynamical solar neighborhood for each cluster based on its mass, to show that action-space tagging is likely still a viable method if contamination from foreground or background stars is unimportant. We summarize and conclude in Section~\ref{sec:conclusion}.

@@

%In Figure~\ref{fig:cluster_offset}, we show the $z$ offset necessary to induce
%the action space dispersion computed from Equation~\ref{eq:action_disp}, using
%the \texttt{gala} computed mass enclosed within a radius of $8.2\,\kpc$ ($\sim
%10^{11}\,M_{\odot}$).

%We now convert the $z$~offset-$m_c$ relation to an $R_{\n}$-$m_c$ relation,
%where $R_{\n}$ indicates the neighborhood radius. An observer
%attempting to reconstruct an open cluster of mass $m_c$ with perfect knowledge
%of the midplane at the location of the cluster would expect to find the
%ejected members using a dynamical cut for field stars within a distance
%$R_{\n}$ of the cluster. If such an observer wants to find ejected stars that
%are at a distance greater than $R_{\n}$, a more sophisticated analysis would
%be required.

>>>>

\shortauthors{Beane et al.}
\shorttitle{The Galactic Midplane is Not a Plane}

@@

\keywords{Galaxy:~disk -- solar~neighborhood --
open~clusters~and~associations:~general -- Galaxy:~structure --
Galaxy:~evolution}

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a revolution in the
era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
local substructure infall \citep{2018MNRAS.481.1501B, 2018arXiv180800451L} or
bar buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation
by a dark matter subtructure in the tidal stream GD1
\citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
indicate that the Milky Way's stellar distribution, which demonstrably departs
significantly from axisymmetry, is undergoing phase mixing and dynamical
interactions across a range of spatial and temporal scales.

@@

@@

In Section~\ref{sec:discussion} we discuss the implications of midplane
variations, and the resulting systematic uncertainty in the vertical action,
for action-space analyses beyond the dynamical solar neighborhood. As an
illustration, in Section \ref{sssec:neighborhood} we use these estimates to
consider the feasibility of one science goal that relies on the invariance of
actions --- the reconstruction of open clusters across large distances using
``dynamical tagging;'' that is, the association of stars to a birth cluster
based on their similar actions. In light of our findings, we considered the
orbits of the nine open clusters within $250\,\pc$ as reported by the {\em
Gaia} collaboration \citep{2018AA...616A..10G}. In
Table~\ref{tab:real_clusters}, we report the distance to each cluster, the
three actions, and the maximum vertical extent of their orbits
($z_{\text{max}}$). In Section~\ref{sssec:reconstruction} we estimate and
discuss the size of the dynamical solar neighborhood for each cluster based on
its mass, to show that action-space tagging is likely still a viable method if
contamination from foreground or background stars is unimportant. We summarize
and conclude in Section~\ref{sec:conclusion}.

@@
